Transferability of clinical prediction models for early trauma care

### 2019-07-03 ManuscriptMarkdown.Rmd
Update ManuscriptMarkdown.Rmd

# Discussion
This study aimed to assess how transfers of clinical prediction models for early
trauma care between different contexts within a single health system affect
mistriage rates. In all data sets, model transfer between samples resulted in
substantial uncertainty in its effect on model performance. Comparing
transferred performance with the reciveing sample validation performance
resulted in **further uncertainty** _(Not very clear what you mean here, please
try to clarify.)_.

@@

## Strengths and limitations
The main strength of this study is that its design realistically reflects
potential model transfers: clinical prediction models for trauma are usually
developed in large metropolitan centres and then transferred as clinical
recommendations to minor non-metropolitan centres. The results should therefore
be of practical importance.

**One limitation is that** _(Try to highlight whether you disucss a strength or
limitation)_ originally all single centres with a valid number of events were to
constitute a data set, but only one centre provided a sufficient number of
events. Leading to the loss of the data set: Individual centres. This data set
might have provided results that could reflect realistic situations as major
trauma centres regularly exchange data.

This study used GCS, RR and SBP as predictors for 30-day
mortality. Additional predictors such as age, mechanism
of injury, or newer predictors as shock index might increase model
performance. Furthermore, we used 30-day mortality as outcome, recognizing
that late mortality, time in hospital, functional impairment, and morbidity are
also important outcome factors in the context of trauma care.

@@

## Clinical applications
This study shows that mistriage can potentially increase following model
transfer, and more prominently in certain cases. This supports the view that
model transfers should be done with caution, as performance might
deteriorate. Recognizing this risks and updating models accordingly should lead
to increased accuracy during trauma patient triage. Potentially saving lives by
minimizing undertriage, and likely save resources by minimizing overtriage. With
registries such as SweTrau, analysis of model transfer could be performed
continuously to optimize models currently in use and advise which model
transfers might be detrimental (or potentially beneficial).

## **Future studies** _(Is this heading required by the journal? If not I suggest we remove these two paragraphs)_
Studies on model transfer can be performed in different settings, e.g. assessing
the transferability between different but similar countries. Also, increasing
the number of patients would increase the power of the results, one way of doing
this would be to replicate this study with the latest data in the SweTrau, as
many hospitals have joined recently leading to an increased number of patients
registered.

Model transfer could potentially be studied using different methods. Staff
opinions considering how applicable this new model is could be obtained using a
questionnaire approach.  Clinical trials using different models could also be
performed, however, appropriate ethical considerations are paramount before
performing this type of study.

## Conclusion
Model transfer resulted in large and varying levels of uncertainty in its effect
on mistriage. Model transfer can therefore be unpredictable, and potentially
lead to increased mistriage, **resulting in increased patient mortality and a
misappropriation of resources, due to increased undertriage and overtriage,
respectively** _(I suggest you move this part to clinical applications and that
you try to express these potential effects a bit more carefully)_. Transfer of
the `r colnames(highest.comp.mistriage)` model lead to an increased mistriage
rate of `r highest.comp.mistriage`. This observed effect was also associated
with considerable uncertainty, meaning mistriage might slightly decrease, but
would more likely increase, potentially substantially. Recognizing the
limitations of this study, the authors believe further studies are warranted due
to the potential economic costs and patient consequences of poor model
transfers.

>>>>

# Discussion
This study aimed to assess how transfers of clinical prediction models for early
trauma care between different contexts within a single health system affect
mistriage rates. In all data sets, model transfer between samples resulted in
substantial uncertainty in its effect on model performance. When comparing
transferred performance to the reciveing sample validation performance the
resulting differance also showed considerable uncertainty.

@@

## Strengths and limitations
### Strenhths
The main strength of this study is that its design realistically reflects
potential model transfers: clinical prediction models for trauma are usually
developed in large metropolitan centres and then transferred as clinical
recommendations to minor non-metropolitan centres. The results should therefore
be of practical importance.

### Limitations
One limitation is that originally all single centres with a valid number of 
events were to constitute a data set, but only one centre provided a sufficient 
number of events. Leading to the loss of the data set: Individual centres. 
This data set might have provided results that could reflect realistic 
situations as major trauma centres regularly exchange data.

Furthermore, this study used GCS, RR and SBP as predictors for 30-day
mortality. Additional predictors such as age, mechanism
of injury, or newer predictors as shock index might increase model
performance. Furthermore, we used 30-day mortality as outcome, recognizing
that late mortality, time in hospital, functional impairment, and morbidity are
also important outcome factors in the context of trauma care.

@@

## Clinical applications
This study shows that mistriage can potentially increase following model
transfer, and more prominently in certain cases. This supports the view that
model transfers should be done with caution, as performance might
deteriorate.

Recognizing this risks and updating models accordingly should lead
to increased accuracy during trauma patient triage. Potentially saving lives by
minimizing undertriage, and likely save resources by minimizing overtriage. With
registries such as SweTrau, analysis of model transfer could be performed
continuously to optimize models currently in use and advise which model
transfers might be detrimental (or potentially beneficial).

## Conclusion
Model transfer resulted in large and varying levels of uncertainty in its effect
on mistriage. Therefore model transfer can be unpredictable, potentially leading 
to increased mistriage, but in some cases also to a decrease in mistriage. 
Noticeably, transfer of the `r colnames(highest.comp.mistriage)` model lead to an 
increased mistriage of `r highest.comp.mistriage`. This observed effect was also 
associated with considerable uncertainty. Recognizing the limitations of this study, 
the authors believe further studies are warranted dueto the potential economic 
costs and patient consequences of poor model transfers.
